\section{System Model}

\begin{figure}[htbp]
\begin{tikzpicture}
    \node[rectangle, draw] (BS) at (0,0) {BS};
    \node[rectangle, draw] (RIS) at (2,1) {RIS: $\bm{\theta}$};
    \node[rectangle, draw] (US) at (4,0) {Users};

    \draw[-{Latex}] (BS.east) -- (RIS.south west) node[midway, above] {$\bm{G}$};
    \draw[-{Latex}] (BS.east) -- (US.west) node[midway, below] {$\bm{h}_{d,k}$};
    \draw[-{Latex}] (RIS.south east) -- (US.west) node[midway, right, above] {$\bm{h}_{r,k}$};
\end{tikzpicture}
\caption{A base concept of the system.}
\end{figure}

The model is described by a downlink MISO system. It consists of three main elements: the Base Station (BS), the users, and the Reconfigurable Intelligent Surface (RIS).
The system has the following parameters:
\begin{itemize}
    \item BS has M antennas.
    \item There are K users to be serviced.
    \item RIS has N elements, which causes a phase shift of $\bm{\theta} \in \mathbb{C}^N$.
    \item $\bm{h}_{d, k} \in \mathbb{C}^{M}$ is the channel matrix between the BS and the user $k$.
    \item $\bm{G} \in \mathbb{C}^{N \text{x} M}$ is the channel matrix between the BS and the RIS.
    \item $\bm{h}_{r, k} \in \mathbb{C}^{N}$ is the channel matrix between the RIS and the user $k$. 
\end{itemize}

Therefore the user $k$ receives the signal:
\begin{equation}
    y_k = \left(\bm{h}_{d,k}^\HH + \bm{h}_{r,k}^\HH \diag(\bm{\theta}) \bm{G}\right) \sum_{i = 1}^{K} \bm{v}_i s_i + n_k
\end{equation}
from the BS.
Where:
\begin{itemize}
    \item $s_i$ is data symbol for user i
    \item $n_k$ add. white Guassian noise. $n_k \sim \mathcal{CN}(0, \sigma^2)$
    \item $\sum_{i = 1}^{K} \bm{v}_i s_i = \bm{x}$ the signal send by the the BS.
    \item $\bm{v}_i \in \mathbb{C}^{M x 1}$ is the beamforming vector from the BS to the user $i$.
    \item $\bm{h}_{d,k}^\HH$ is direct path between the BS and user $k$.
    \item $\bm{h}_{r,k}^\HH \diag(\bm{\theta}) \bm{G}$ is the path from the BS to RIS, application of the phase shift and then the path form the RIS to the user $k$.
    \item Summary: $y_k = (\bm{h}_{dir}^\HH + \bm{h}_{indir}^\HH) \bm{x} + n_k$
\end{itemize}

The SINR can be calculated as:
\begin{equation}
    \alpha_k = \frac{\left|\left(\bm{h}_{d,k}^\HH + \bm{h}_{r,k}^\HH \diag(\bm{\theta}) \bm{G}\bm{v}_k\right)\right|^2}{\sum_{i=1, i \ne k}^{K}\left|\left(\bm{h}_{d,k}^\HH + \bm{h}_{r,k}^\HH \diag(\bm{\theta}) \bm{G}\bm{v}_i\right)\right|^2 + \sigma_0^2}
\end{equation}

which is simply the expected signal divided by the sum of the noise and all other signals sent.


The objective is to maximize the weighted sum rate (WSR):

\begin{align}
    (\mathcal{P}_1): \quad 
    \max_{\mathbf{V}, \bm{\theta}} \quad 
    & \sum_{k=1}^{K} \omega_k \log(1 + \alpha_k) \\
    \text{s.t.} \quad 
    & |\theta_n| = 1, \ \forall n, \label{eq:theta_constraint} \\
    & \sum_{k=1}^{K} \|\bm{v}_k\|^2 \le P_{\max}
\end{align}

The system allows changing the beamforming vectors of the BS and the phase shift values of the RIS elements.
$\bm{V} = [\bm{v}_1, ..., \bm{v}_K]$ and $\omega_k$ a weight coefficient for the user $k$.

This is a typical multi-ratio fractional programming problem, with a possibly non-convex solution space.